---
title: "Graphic Design with ggplot2"
subtitle: "<span style='font-size:1.75em;line-height:1.1;'>Concepts of the <code>{ggplot2}</code> Package Pt. 2:<br><span style='font-weight:500;'>Solution Exercise 1</span></span>"
author: "<b style='color:#28a87d;'>Cédric Scherer</b> **//** rstudio::conf **//** July 2022"
format:
  revealjs:
    theme: css/style.scss
    footer: '**[Cédric Scherer](https::/cedricscherer.com) //** rstudio::conf **//** July 2022'
    #'rstudio::conf(workshop = "Graphic Design with ggplot2", instructor = "Cédric Scherer")'
    transition: fade
    progress: true
    multiplex: false
    preview-links: true
    scrollable: false
    hide-inactive-cursor: true
    highlight-style: printing
    chalkboard:
      theme: whiteboard
      boardmarker-width: 4
      buttons: true
    pause: true
    mainfont: Cabinet Grotesk
editor: source
knitr:
  opts_chunk:
    dev: "ragg_png"
    retina: 1
    dpi: 200
execute:
  freeze: auto
  cache: true
  echo: true
  fig-width: 5
  fig-height: 6
---

## Exercise 1

::: question
-   Have a look at the following visualization of the cumulative time that cosmo- and astronauts have spent in outer space. The data also contains information on the year of their first and last travel, respectively.
-   Together with your group, discuss which layers and modifications are needed to create such a chart with `{ggplot2}`.
    -   Note down the aesthetics, geometries, and scales used for each element of this graphic.
    -   What is the coordinate system? Have any adjustments been made?
    -   Which theme was used and how was it modified?
:::

------------------------------------------------------------------------

### Layers

::: incremental
-   **`geom_point()`**
    -   `aes(x = id, y = hours, size = hours)`
-   **`geom_linerange()`**
    -   `aes(x = id, ymin = 0, ymax = hours, color = hours, alpha = hours)`
-   **`geom_point()`**
    -   `aes(x = id, y = 0), shape = 15, color = "#808080"`
-   **`geom_text()`**
    -   `aes(x = id, y = 0, label = year), size = 4.5, hjust = 1.2`
-   **`geom_text()`**
    -   `aes(x = id, y = hours, label = max), size = 3.9, vjust = -.35`
:::

------------------------------------------------------------------------

### Scales

::: incremental
-   **`scale_x_continuous()`**
    -   `limits = c(-300, NA), expand = c(0, 0)`
-   **`scale_y_continuous()`**
    -   `limits = c(0, 230000), expand = c(0, 0)`
-   **`scale_color_distiller()`**
    -   `palette = "YlGnBu, direction = -1`
-   **`scale_size()`**
    -   `range = c(.001, 3)`
-   **`scale_alpha()`**
    -   `range = c(.33, .95)`
:::

------------------------------------------------------------------------

### Coordinate System

::: incremental
-   **`coord_polar()`**
    -   `theta = "y"`
:::

------------------------------------------------------------------------

### Coordinate System

-   **`coord_polar()`**
    -   `theta = "y"`

### Theme

::: incremental
-   **`theme_void()`**
    -   `legend.position = "none"`
    -   `plot.background = element_rect(fill = "black")`
    -   `plot.margin = margin(-70, -70, -70, -70)`
    -   `plot.caption = element_text(hjust = .5, margin = margin(-100, 0, 100, 0), ...)`
:::

------------------------------------------------------------------------

### Coordinate System

-   **`coord_polar()`**
    -   `theta = "y"`

### Theme

-   **`theme_void()`**
    -   `legend.position = "none"`
    -   `plot.background = element_rect(fill = "black")`
    -   `plot.margin = margin(-70, -70, -70, -70)`
    -   `plot.caption = element_text(...)`

### Title

::: incremental
-   **2 x `annotate(geom = "text", x = -300, y = 0, ...)`**
:::

## Data Prep

```{r}
#| label: import-data-astronauts
library(tidyverse)

df_astro <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-14/astronauts.csv'
)

df_missions <-
  df_astro %>%
  group_by(name) %>%
  summarize(
    hours = sum(hours_mission),
    year = min(year_of_mission),
    max_year = max(year_of_mission)
  ) %>%
  ungroup() %>%
  mutate(year = -year) %>%
  arrange(year) %>%
  mutate(id = row_number())

```

## Code Pt. 1

```{r}
#| label: astronout-viz-blank
# install.packages("scico")

g1 <-
  ggplot(df_missions, aes(x = id, y = hours, color = hours)
    ) +
    ## curves
    geom_linerange(aes(ymin = 0, ymax = hours, alpha = hours), size = .25) +
    ## baseline
    geom_point(aes(y = 0), shape = 15, size = .1, color = "#808080") +
    ## points
    geom_point(aes(y = hours, size = hours)) +
    ## turn into circular
    coord_polar(theta = "y", start = 0, clip = "off") +
    ## add axis spacings
    scale_x_continuous(limits = c(-300, NA), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 23000), expand = c(0, 0)) +
    ## change colors, transparencies, and bubble sizes
    scale_color_distiller(palette = "YlGnBu", direction = -1) +
    scale_size(range = c(.001, 3)) +
    scale_alpha(range = c(.33, .95)) +
    ## remove all theme components
    theme_void() +
    theme(
      ## set dark background
      plot.background = element_rect(fill = "black"),
      ## remove "white" space
      plot.margin = margin(-70, -70, -70, -70),
      ## remove legends
      legend.position = "none"
    )
```

## Data Prep Labels

```{r}
#| label: astronout-viz-label-dfs
df_labs <-
  df_missions %>%
  filter(year %in% -c(1961, 197:201*10, 2019)) %>%
  group_by(year) %>%
  filter(id == min(id))

df_max <-
  df_missions %>%
  arrange(-hours) %>%
  slice(1) %>%
  mutate(
    first_name = str_remove(name, ".*, "),
    last_name = str_remove(name, "(?<=),.*"),
    label = paste("Between", abs(year), "and", max_year, ",\n", first_name, last_name, "has spent\n", format(hours, big.mark = ','), "hours in space.\nThat's roughly", round(hours / 24, 0), "days!")
  )
```

## Code Pt. 2

```{r}
#| label: astronout-viz-labels
g2 <-
  g1 +
    ## labels years
    geom_text(
      data = df_labs, aes(y = 0, label = abs(year)),
      family = "Lato", fontface = "bold", color = "#808080",
      size = 4.5, hjust = 1.2
    ) +
    ## label max
    geom_text(
      data = df_max, aes(label = label),
      family = "Lato", size = 3.9, vjust = -.35
    ) +
    ## title shadow
    annotate(
      geom = "text", x = -300, y = 0, label = "Travelling to\nOuter Space",
      family = "Boska", fontface = "bold", lineheight = .9,
      size = 20, color = "white", hjust = .57, vjust = .45, alpha = .25
    ) +
    ## title
    annotate(
      geom = "text", x = -300, y = 0, label = "Travelling to\nOuter Space",
      family = "Boska", fontface = "bold", lineheight = .85,
      size = 20, color = "white", hjust = .55, vjust = .4
    ) +
    ## caption
    labs(caption = "Cumulative time in outer space for all 565 cosmonauts and astronauts who participated
    in space missions between April 23, 1961 and January 15, 2020, sorted by the year of their first mission.

    Graphic: Cédric Scherer • Data: Stavnichuk & Corlett 2020 (DOI: 10.17632/86tsnnbv2w.1)") +
    ## modify caption + move inside plot area
    theme(
      plot.caption = element_text(
        family = "Lato",
        size = 15, color = "#808080", hjust = .5,
        margin = margin(-100, 0, 100, 0)
      )
    )
```

------------------------------------------------------------------------

```{r}
#| label: astronout-viz-plot
#| echo: false
#| fig-width: 15
#| fig-height: 16
#| fig-align: "center"
g2
```

## Code with Special Extensions

```{r}
#| label: astronout-viz-extensions
#| code-line-numbers: "1,2,3,7,8,10,11,24,26"
# install.packages("ggforce")
# install.packages("scico")
# devtools::install_github("coolbutuseless/ggblur")

g_ext <-
  ggplot(df_missions, aes(x = id, y = hours, color = hours)) +
    ## geom_link() from {ggforce} to draw smooth curves
    ggforce::geom_link(aes(xend = id, yend = 0, alpha = hours), size = .25, n = 300) +
    geom_point(aes(y = 0), shape = 15, size = .1, color = "#808080") +
    ##geom_point_blur() from {ggblur} to add points with gradual fading
    ggblur::geom_point_blur(aes(size = hours, blur_size = hours), blur_steps = 25) +
    geom_text(
      data = df_labs, aes(y = 0, label = abs(year)),
      family = "Lato", fontface = "bold", color = "#808080",
      size = 4.5, hjust = 1.2
    ) +
    geom_text(
      data = df_max, aes(label = label),
      family = "Lato", size = 3.9, vjust = -.35
    ) +
    coord_polar(theta = "y", start = 0, clip = "off") +
    scale_x_continuous(limits = c(-300, NA), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 23000), expand = c(0, 0)) +
    ## use custom color palette from {scico}
    scico::scale_color_scico(palette = "buda") +
    scale_size(range = c(.001, 3)) +
    ggblur::scale_blur_size_continuous(range = c(.5, 10), guide = "none") +
    scale_alpha(range = c(.33, .95)) +
    annotate(
      geom = "text", x = -300, y = 0, label = "Travelling to\nOuter Space",
      family = "Boska", fontface = "bold", lineheight = .9,
      size = 20, color = "white", hjust = .57, vjust = .45, alpha = .25
    ) +
    annotate(
      geom = "text", x = -300, y = 0, label = "Travelling to\nOuter Space",
      family = "Boska", fontface = "bold", lineheight = .85,
      size = 20, color = "white", hjust = .55, vjust = .4
    ) +
    labs(caption = "Cumulative time in outer space for all 565 cosmonauts and astronauts who participated
    in space missions between April 23, 1961 and January 15, 2020, sorted by the year of their first mission.

    Graphic: Cédric Scherer • Data: Stavnichuk & Corlett 2020 (DOI: 10.17632/86tsnnbv2w.1)") +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black"),
      plot.margin = margin(-70, -70, -70, -70),
      legend.position = "none",
      plot.caption = element_text(
        family = "Lato",
        size = 15, color = "#808080", hjust = .5,
        margin = margin(-100, 0, 100, 0)
      )
    )
```

------------------------------------------------------------------------

```{r}
#| label: astronout-viz-extensions-plot
#| echo: false
#| fig-width: 15
#| fig-height: 16
#| fig-align: "center"
g_ext
```
