---
title: "Graphic Design with ggplot2"
subtitle: "<span style='font-size:1.75em;line-height:1.1;'>Concepts of the <code>{ggplot2}</code> Package Pt. 2:<br><span style='font-weight:500;'>Solution Exercise 2</span></span>"
author: "<b style='color:#28a87d;'><a href='https://www.cedricscherer.com'>Cédric Scherer</a></b> **//** rstudio::conf **//** July 2022"
format:
  revealjs:
    theme: css/style.scss
    footer: '**[Cédric Scherer](https::/cedricscherer.com) //** rstudio::conf **//** July 2022'
    #'rstudio::conf(workshop = "Graphic Design with ggplot2", instructor = "Cédric Scherer")'
    transition: fade
    progress: true
    multiplex: false
    preview-links: true
    scrollable: false
    hide-inactive-cursor: true
    highlight-style: printing
    chalkboard:
      theme: whiteboard
      boardmarker-width: 4
      buttons: true
    pause: true
    mainfont: Cabinet Grotesk
editor: source
knitr:
  opts_chunk:
    dev: "ragg_png"
    retina: 1
    dpi: 200
execute:
  freeze: auto
  cache: true
  echo: true
  fig-width: 5
  fig-height: 6
---


## Exercise 2

::: question
* Create a circular lollipop plot of reported bike shares per season.
  - The data is not in the right format as there are no totals.<br>How can you solve it?
  - Remove all legend elements (with a single line of code).
  - How can you add the labels next to the starting point of each lollipop?
  - How could you add a baseline?
:::


------------------------------------------------------------------------


```{r}
#| label: ciruclar-lollipop-exercise
#| echo: false
#| fig-width: 6
#| fig-height: 6
#| fig-align: "center"
library(tidyverse)

bikes <- readr::read_csv(
  here::here("data", "london-bikes-custom.csv"),
  col_types = "Dcfffilllddddc"
)

bikes$season <- forcats::fct_inorder(bikes$season)

theme_set(theme_light(base_size = 14, base_family = "Roboto Condensed"))

ggplot(bikes, aes(x = as.numeric(season), y = count, group = season)) +
  stat_summary(geom = "point", fun = "sum", size = 3) +
  stat_summary(geom = "linerange", ymin = 0, fun.max = function(y) sum(y)) +
  stat_summary(
    geom = "text", aes(label = season, y = 0),
    family = "Cabinet Grotesk", fontface = "bold", size = 4.5, hjust = 1.15
  ) +
  annotate(
    geom = "linerange", xmin = .7, xmax = 4.3, y = 0
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(expand = c(.5, .5)) +
  scale_y_continuous(limits = c(0, 7.5*10^6)) +
  theme_void() +
  theme(plot.margin = margin(rep(-130, 4)))
```



## Preparation

```{r}
#| label: prep
library(tidyverse)

bikes <- readr::read_csv(
  here::here("data", "london-bikes-custom.csv"),
  col_types = "Dcfffilllddddc"
)

bikes$season <- stringr::str_to_title(bikes$season)
bikes$season <- forcats::fct_inorder(bikes$season)

theme_set(theme_light(base_size = 14, base_family = "Roboto Condensed"))
```



## Lollipop Plot with Pre-Calculated Sums

```{r}
#| label: concepts-exercise-lollipop-cartesian-data
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count))
```

## Lollipop Plot with Pre-Calculated Sums

```{r}
#| label: concepts-pt2-exercise-2-1
#| output-location: column
#| code-line-numbers: "1,2,3|4,5,6,7,8"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  )
```



## Calculate Sums via `stat_summary()`

```{r}
#| label: concepts-exercise-lollipop-cartesian-stat
#| output-location: column
#| code-line-numbers: "2,3,4|5,6,7,8"
ggplot(bikes, aes(x = season, y = count)) +
  stat_summary(
    geom = "point", fun = "sum", size = 3
  ) +
  stat_summary(
    geom = "linerange", ymin = 0,
    fun.max = function(y) sum(y)
  )
```



## Calculate Sums via `stat_summary()`

```{r}
#| label: concepts-exercise-lollipop-cartesian-stat-zero
#| output-location: column
#| code-line-numbers: "9"
ggplot(bikes, aes(x = season, y = count)) +
  stat_summary(
    geom = "point", fun = "sum", size = 3
  ) +
  stat_summary(
    geom = "linerange", ymin = 0,
    fun.max = function(y) sum(y)
  ) +
  coord_cartesian(ylim = c(0, NA))
```



## Apply a Polar Coordinate System

```{r}
#| label: concepts-pt2-exercise-2-2
#| output-location: column
#| code-line-numbers: "9"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  coord_polar(theta = "y")
```



## Fix Axis Ranges

```{r}
#| label: concepts-pt2-exercise-2-3
#| output-location: column
#| code-line-numbers: "10,11,12,13,14,15"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  )
```



## Remove All Theme Components

```{r}
#| label: concepts-pt2-exercise-2-4
#| output-location: column
#| code-line-numbers: "16"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void()
```



## Fix Plot Margin

```{r}
#| label: concepts-pt2-exercise-2-5
#| output-location: column
#| code-line-numbers: "17"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```


## Add "Axis" Text

```{r}
#| label: concepts-pt2-exercise-2-6
#| output-location: column
#| code-line-numbers: "9,10,11"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  geom_text(
    aes(label = season, y = 0)
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```



## Style "Axis" Text

```{r}
#| label: concepts-pt2-exercise-2-7
#| output-location: column
#| code-line-numbers: "11,12"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  geom_text(
    aes(label = season, y = 0),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```



## Alternatively: Add Direct Labels

```{r}
#| label: concepts-exercise-lollipop-label-season-direct
#| output-location: column
#| code-line-numbers: "14,15,16,17"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  geom_text(
    aes(label = season, y = 0),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  geom_text(
    aes(label = paste0(round(count / 10^6, 1), "M")),
    size = 4, vjust = -1, family = "Cabinet Grotesk"
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```



## Add a Baseline — *ugly but simple*

```{r}
#| label: concepts-exercise-lollipop-baseline-geom-hline
#| output-location: column
#| code-line-numbers: "9"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  geom_hline(yintercept = 0) +
  geom_text(
    aes(label = season, y = 0),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```



## Add a Baseline — *nice but unusual*

```{r}
#| label: concepts-exercise-lollipop-baseline-geom-linerange
#| output-location: column
#| code-line-numbers: "4|4,9,10,11"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  geom_linerange(
    xmin = .7, xmax = 4.3, y = 0
  ) +
  geom_text(
    aes(label = season, y = 0),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```



## Add a Baseline — *yeah, that's it!*

```{r}
#| label: concepts-pt2-exercise-2-8
#| output-location: column
#| code-line-numbers: "4,9,10,11,12"
bikes %>%
  group_by(season) %>%
  summarize(count = sum(count)) %>%
  ggplot(aes(x = season, y = count)) +
  geom_point(size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = count)
  ) +
  annotate(
    geom = "linerange",
    xmin = .7, xmax = 4.3, y = 0
  )  +
  geom_text(
    aes(label = season, y = 0),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```


## Solution using `stat_summary()`

```{r}
#| label: concepts-exercise-lollipop-stat-summary
#| output-location: column
#| code-line-numbers: "1|2,3,4|5,6,7,8|9,10,11,12,13,14"
ggplot(bikes, aes(x = as.numeric(season), y = count)) +
  stat_summary(
    geom = "point", fun = "sum", size = 3
  ) +
  stat_summary(
    geom = "linerange", ymin = 0,
    fun.max = function(y) sum(y)
  ) +
  stat_summary(
    geom = "text",
    aes(
      label = season,
      y = 0
    ),
    family = "Cabinet Grotesk", size = 4.5,
    fontface = "bold", hjust = 1.15
  ) +
  annotate(
    geom = "linerange",
    xmin = .7, xmax = 4.3, y = 0
  ) +
  coord_polar(theta = "y") +
  scale_x_discrete(
    expand = c(.5, .5)
  ) +
  scale_y_continuous(
    limits = c(0, 7.5*10^6)
  ) +
  theme_void() +
  theme(plot.margin = margin(rep(-100, 4)))
```
