---
title: "Graphic Design with ggplot2"
subtitle: "<span style='font-size:1.75em;line-height:1.1;'>Working with Labels and Annotations:<br><span style='font-weight:500;'>Solution Exercise 1</span></span>"
author: "<b style='color:#28a87d;'>Cédric Scherer</b> **//** rstudio::conf **//** July 2022"
format: 
  revealjs: 
    theme: css/style.scss
    footer: '**[Cédric Scherer](https::/cedricscherer.com) //** rstudio::conf **//** July 2022'
    #'rstudio::conf(workshop = "Graphic Design with ggplot2", instructor = "Cédric Scherer")'
    transition: fade
    progress: true
    multiplex: false
    preview-links: true
    scrollable: false
    hide-inactive-cursor: true
    highlight-style: printing
    chalkboard:
      theme: whiteboard
      boardmarker-width: 4
      buttons: true
    pause: true
    mainfont: Cabinet Grotesk
editor: source
knitr:
  opts_chunk: 
    dev: "ragg_png"
    retina: 1
    dpi: 300
execute:
  freeze: auto
  echo: true
  fig-width: 5
  fig-height: 6
---


## Exercise 1

::: question
- {ggtext} also comes with some new geom's. Explore those and other options on the package webpage: [wilkelab.rg/ggtext](https://wilkelab.org/ggtext/).
- **Create the following visualization, as close as possible**, with the `penguins` dataset which is provided by the [{palmerpenguins} package](https://allisonhorst.github.io/palmerpenguins/).
  - For the species labels, you likely have to create a summary data set.
  - Use the {ggtext} geometries and theme elements to format the labels.
  - Also, make use of the other components such as scales, original theme, and theme customization.
:::


------------------------------------------------------------------------


```{r}
#| label: exercise-text-styling
#| fig-width: 6
#| fig-height: 3.7
#| fig-align: "center"
#| echo: false
#| purl: false
library(tidyverse)
library(palmerpenguins)

penguins_labs <-
  penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(
    species_lab = case_when(
      species == "Adelie" ~ "<b style='font-size:15pt;'>*P. adéliae*</b><br>(Adélie penguin)",
      species == "Chinstrap" ~ "<b style='font-size:15pt;'>*P. antarctica*</b><br>(Chinstrap penguin)",
      species == "Gentoo" ~ "<b style='font-size:15pt;'>*P. papua*</b><br>(Gentoo penguin)"
    )
  )

ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2, stroke = .3) +
  geom_point(shape = 1, stroke = .3) +
  ggtext::geom_richtext(
    data = penguins_labs,
    aes(label = species_lab, color = species, 
        color = after_scale(prismatic::clr_darken(color, .4))),
    family = "Roboto Condensed",
    size = 3, lineheight = .8,
    fill = "#ffffffab", ## hex-alpha code
    show.legend = FALSE
  ) +
  coord_cartesian(
    expand = FALSE,
    clip = "off"
  )  +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5)
  ) +
  scale_color_manual(
    guide = "none",
    values = c("#FF8C00", "#A034F0", "#159090")
  ) +
  scale_size(
    name = "Body mass:",
    breaks = 3:6 * 1000,
    labels = function(x) paste(x / 1000, "kg"),
    range = c(.25, 4.5)
  ) +
  labs(
    x = "Bill length *(mm)*",
    y = "Bill depth *(mm)*",
    title = "Bill dimensions of brush-tailed penguins *Pygoscelis spec.*",
    caption = "Horst AM, Hill AP, Gorman KB (2020). <span style='font-family:tabular;'>palmerpenguins</span> R package version 0.1.0"
  ) +
  theme_minimal(
    base_size = 10, base_family = "Roboto Condensed"
  ) +
  theme(
    plot.title = ggtext::element_markdown(
      face = "bold", size = 16, margin = margin(12, 0, 12, 0)
    ),
    plot.title.position = "plot",
    plot.caption = ggtext::element_markdown(
      size = 7, color = "grey50",
      margin = margin(12, 0, 6, 0)
    ),
    plot.caption.position = "plot",
    axis.text = element_text(family = "Tabular"),
    axis.title.x = ggtext::element_markdown(margin = margin(t = 8)),
    axis.title.y = ggtext::element_markdown(margin = margin(r = 8)),
    panel.grid.minor = element_blank(),
    legend.text = element_text(color = "grey50"),
    plot.margin = margin(0, 14, 0, 12),
    plot.background = element_rect(fill = NA, color = "grey50", size = 1)
  )
```




## The Data Set

```{r}
#| label: prep
# install.packages("palmerpenguins")
library(palmerpenguins)

penguins
```



## Create a Labeled Bubble Chart

```{r}
#| label: annotations-exercise-1-1
#| eval: false
#| code-line-numbers: "1,2,3,4,5,6|7,8,9,10,11,12"
ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2) +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    title = "Bill dimensions of brush-tailed penguins Pygoscelis spec.",
    caption = "Horst AM, Hill AP, Gorman KB (2020). palmerpenguins R package version 0.1.0"
  )
```

## A Labelled Bubble Plot

```{r}
#| label: annotations-exercise-1-1
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Add a Custom Theme

```{r}
#| label: annotations-exercise-1-2
#| eval: false
#| code-line-numbers: "13,14,15,16,17|13|14,15,16,17"
ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2) +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    title = "Bill dimensions of brush-tailed penguins Pygoscelis spec.",
    caption = "Horst AM, Hill AP, Gorman KB (2020). palmerpenguins R package version 0.1.0"
  ) +
  theme_minimal(base_size = 10, base_family = "Roboto Condensed") +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot",
    panel.grid.minor = element_blank()
  )
```

## Add a Custom Theme

```{r}
#| label: annotations-exercise-1-2
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Add an Outline

```{r}
#| label: annotations-exercise-1-3
#| eval: false
#| code-line-numbers: "8"
p1 <- 
  ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2, stroke = .3) +
  geom_point(shape = 1, stroke = .3) +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    title = "Bill dimensions of brush-tailed penguins Pygoscelis spec.",
    caption = "Horst AM, Hill AP, Gorman KB (2020). palmerpenguins R package version 0.1.0"
  ) +
  theme_minimal(base_size = 10, base_family = "Roboto Condensed") +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot",
    panel.grid.minor = element_blank()
  )

p1
```

## Add an Outline

```{r}
#| label: annotations-exercise-1-3
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Style Color Legend

```{r}
#| label: annotations-exercise-1-4
#| eval: false
#| code-line-numbers: "2,3,4,5"
p1 +
  scale_color_manual(
    name = "Species:",
    values = c("#FF8C00", "#A034F0", "#159090")
  )
```

## Style Color Legend

```{r}
#| label: annotations-exercise-1-4
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Style Size Legend

```{r}
#| label: annotations-exercise-1-5
#| eval: false
#| code-line-numbers: "13,14,15,16,17,18|13,17|16"
p2 <- p1 +
  scale_color_manual(
    name = "Species:",
    values = c("#FF8C00", "#A034F0", "#159090")
  ) +
  scale_size(
    name = "Body mass:",
    breaks = 3:6 * 1000,
    labels = function(x) paste(x / 1000, "kg"),
    range = c(.5, 5)
  )

p2
```

## Style Size Legend

```{r}
#| label: annotations-exercise-1-5
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Adjust Axes

```{r}
#| label: annotations-exercise-1-6
#| eval: false
#| code-line-numbers: "2,3,4,5,6,7,8,9,10,11"
p2 +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5),
    expand = c(0, 0)
  ) 
```

## Adjust Axes

```{r}
#| label: annotations-exercise-1-6
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Adjust Axes

```{r}
#| label: annotations-exercise-1-7
#| eval: false
#| code-line-numbers: "2,3,4,5"
p3 <- p2 +
  coord_cartesian(
    expand = FALSE,
    clip = "off"
  ) +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5)
  )

p3
```

## Adjust Axes

```{r}
#| label: annotations-exercise-1-7
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Fixed Coordinate System?

```{r}
#| label: annotations-exercise-1-7-fixed
#| eval: false
#| code-line-numbers: "2"
p2 +
  coord_fixed(
    expand = FALSE,
    clip = "off"
  ) +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5)
  )
```

## Fixed Coordinate System?

```{r}
#| label: annotations-exercise-1-7-fixed
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Format Labels with {ggtext}

```{r}
#| label: annotations-exercise-1-8
#| eval: false
#| code-line-numbers: "3,4,5,6"
p3 +
  labs(
    x = "Bill length *(mm)*",
    y = "Bill depth *(mm)*",
    title = "Bill dimensions of brush-tailed penguins *Pygoscelis spec.*",
    caption = "Horst AM, Hill AP, Gorman KB (2020). <span style='font-family:tabular;'>palmerpenguins</span> R package version 0.1.0"
  ) 
```

## Format Labels with {ggtext}

```{r}
#| label: annotations-exercise-1-8
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Format Labels with {ggtext}

```{r}
#| label: annotations-exercise-1-9
#| eval: false
#| code-line-numbers: "1,11,13,14,16,17,18|12,15"
library(ggtext)

p4 <- p3 +
  labs(
    x = "Bill length *(mm)*",
    y = "Bill depth *(mm)*",
    title = "Bill dimensions of brush-tailed penguins *Pygoscelis spec.*",
    caption = "Horst AM, Hill AP, Gorman KB (2020). <span style='font-family:tabular;'>palmerpenguins</span> R package version 0.1.0"
  ) +
  theme(
    plot.title = element_markdown(
      face = "bold", size = 16, margin = margin(12, 0, 12, 0)
    ),
    plot.caption = element_markdown(
      size = 7, color = "grey50", margin = margin(12, 0, 6, 0)
    ),
    axis.title.x = element_markdown(margin = margin(t = 8)),
    axis.title.y = element_markdown(margin = margin(r = 8))
  )

p4
```

## Format Labels with {ggtext}

```{r}
#| label: annotations-exercise-1-9
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Style Other Theme Elements

```{r}
#| label: annotations-exercise-1-10
#| eval: false
#| code-line-numbers: "2,3,4,5,6,7|3|4|5,6"
p5 <- p4 + 
  theme(
    axis.text = element_text(family = "Tabular"),
    legend.text = element_text(color = "grey50"),
    plot.margin = margin(0, 14, 0, 12),
    plot.background = element_rect(fill = NA, color = "grey50", size = 1)
  )

p5
```


## Style Other Theme Elements

```{r}
#| label: annotations-exercise-1-10
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```




## Create the Summary Data

```{r}
#| label: summary-data
#| code-line-numbers: "1,2,3,4,5,6,7,8,9,10,11,12,13|3|4,5|5,6|7,8,9,10,11,12,13|8|9"
library(tidyverse)

penguins_labs <-
  penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(
    species_lab = case_when(
      species == "Adelie" ~ "<b style='font-size:15pt;'>*P. adéliae*</b><br>(Adélie penguin)",
      species == "Chinstrap" ~ "<b style='font-size:15pt;'>*P. antarctica*</b><br>(Chinstrap penguin)",
      species == "Gentoo" ~ "<b style='font-size:15pt;'>*P. papua*</b><br>(Gentoo penguin)"
    )
  )

penguins_labs
```



## Add Species Annotations

```{r}
#| label: annotations-exercise-1-11
#| eval: false
#| code-line-numbers: "2,3,4,5,6|3|4,5"
p5 +
  geom_richtext(
    data = penguins_labs,
    aes(label = species_lab),
    color = "black", size = 3
  )
```

## Add Species Annotations

```{r}
#| label: annotations-exercise-1-11
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Style Species Annotations

```{r}
#| label: annotations-exercise-1-12
#| eval: false
#| code-line-numbers: "6,7,8,9|6,7|8|9"
p5 +
  geom_richtext(
    data = penguins_labs,
    aes(label = species_lab),
    color = "black", size = 3,
    family = "Roboto Condensed",
    lineheight = .8,
    fill = "#ffffffab", ## hex-alpha code
    show.legend = FALSE
  )
```

## Style Species Annotations

```{r}
#| label: annotations-exercise-1-12
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Style Species Annotations

```{r}
#| label: annotations-exercise-1-13
#| eval: false
#| code-line-numbers: "5,6|6"
p5 +
  geom_richtext(
    data = penguins_labs,
    aes(label = species_lab, 
        color = species, 
        color = after_scale(prismatic::clr_darken(color, .4))),
    family = "Roboto Condensed",
    size = 3, lineheight = .8,
    fill = "#ffffffab", 
    show.legend = FALSE
  )
```

## Style Species Annotations

```{r}
#| label: annotations-exercise-1-13
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## ... and Remove Color Legend

```{r}
#| label: annotations-exercise-1-14
#| eval: false
#| code-line-numbers: "13"
p5 +
  geom_richtext(
    data = penguins_labs,
    aes(label = species_lab, 
        color = species, 
        color = after_scale(prismatic::clr_darken(color, .4))),
    family = "Roboto Condensed",
    size = 3, lineheight = .8,
    fill = "#ffffffab", ## hex-alpha code
    show.legend = FALSE
  ) +
  scale_color_manual(
    guide = "none",
    values = c("#FF8C00", "#A034F0", "#159090")
  )
```



## ... and Remove Color Legend

```{r}
#| label: annotations-exercise-1-14
#| echo: false
#| fig-width: 6
#| fig-height: 3.7
```



## Full Code

```{r}
#| label: annotations-exercise-full-code
#| eval: false
library(tidyverse)
library(palmerpenguins)
library(ggtext)

penguins_labs <-
  penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(
    species_lab = case_when(
      species == "Adelie" ~ "<b style='font-size:15pt;'>*P. adéliae*</b><br>(Adélie penguin)",
      species == "Chinstrap" ~ "<b style='font-size:15pt;'>*P. antarctica*</b><br>(Chinstrap penguin)",
      species == "Gentoo" ~ "<b style='font-size:15pt;'>*P. papua*</b><br>(Gentoo penguin)"
    )
  )

ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2, stroke = .3) +
  geom_point(shape = 1, stroke = .3) +
  geom_richtext(
    data = penguins_labs,
    aes(label = species_lab, 
        color = species, 
        color = after_scale(prismatic::clr_darken(color, .4))),
    family = "Roboto Condensed",
    size = 3, lineheight = .8,
    fill = "#ffffffab", ## hex-alpha code
    show.legend = FALSE
  ) +
  coord_cartesian(
    expand = FALSE,
    clip = "off"
  )  +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5)
  ) +
  scale_color_manual(
    guide = "none",
    values = c("#FF8C00", "#A034F0", "#159090")
  ) +
  scale_size(
    name = "Body mass:",
    breaks = 3:6 * 1000,
    labels = function(x) paste(x / 1000, "kg"),
    range = c(.25, 4.5)
  ) +
  labs(
    x = "Bill length *(mm)*",
    y = "Bill depth *(mm)*",
    title = "Bill dimensions of brush-tailed penguins *Pygoscelis spec.*",
    caption = "Horst AM, Hill AP, Gorman KB (2020). <span style='font-family:tabular;'>palmerpenguins</span> R package version 0.1.0"
  ) +
  theme_minimal(
    base_size = 10, base_family = "Roboto Condensed"
  ) +
  theme(
    plot.title = element_markdown(
      face = "bold", size = 16, margin = margin(12, 0, 12, 0)
    ),
    plot.title.position = "plot",
    plot.caption = element_markdown(
      size = 7, color = "grey50",
      margin = margin(12, 0, 6, 0)
    ),
    plot.caption.position = "plot",
    axis.text = element_text(family = "Tabular"),
    axis.title.x = element_markdown(margin = margin(t = 8)),
    axis.title.y = element_markdown(margin = margin(r = 8)),
    panel.grid.minor = element_blank(),
    legend.text = element_text(color = "grey50"),
    plot.margin = margin(0, 14, 0, 12),
    plot.background = element_rect(fill = NA, color = "grey50", size = 1)
  )
```



