---
title: "Graphic Design with ggplot2"
subtitle: "<span style='font-size:1.75em;line-height:1.1;'>Working with Colors:<br><span style='font-weight:500;'>Solution Exercise 1</span></span>"
author: "<b style='color:#28a87d;'>Cédric Scherer</b> **//** rstudio::conf **//** July 2022"
format: 
  revealjs: 
    theme: css/style.scss
    footer: '**[Cédric Scherer](https::/cedricscherer.com) //** rstudio::conf **//** July 2022'
    #'rstudio::conf(workshop = "Graphic Design with ggplot2", instructor = "Cédric Scherer")'
    transition: fade
    progress: true
    multiplex: false
    preview-links: true
    scrollable: false
    hide-inactive-cursor: true
    highlight-style: printing
    chalkboard:
      theme: whiteboard
      boardmarker-width: 4
      buttons: true
    pause: true
    mainfont: Cabinet Grotesk
editor: source
knitr:
  opts_chunk: 
    dev: "ragg_png"
    retina: 1
    dpi: 300
execute:
  freeze: auto
  echo: true
  fig-width: 5
  fig-height: 6
---

## Exercise

::: question
-   **Create a similar visualization as close as possible:**
:::

```{r}
#| label: exercise-jittered-errorbars
#| echo: false
#| fig-width: 10
#| fig-align: "center"
library(tidyverse)
library(colorspace)
library(ggtext)

bikes <- readr::read_csv(
  "https://raw.githubusercontent.com/z3tt/graphic-design-ggplot2/main/data/london-bikes-custom.csv", 
  col_types = "Dcfffilllddddc"
)

bikes$season <- forcats::fct_inorder(bikes$season)

colors <- c("#987708", "#663399")

ggplot(bikes, aes(x = season, y = count)) +
  ggforce::geom_sina(
    aes(
      color = stage(
        day_night, after_scale = lighten(color, .6)
    )),
    position = position_dodge(width = .6),
    alpha = .5
  ) + 
  stat_summary(
    aes(color = day_night),
    position = position_dodge(width = .6),
    fun = mean,
    fun.max = function(y) mean(y) + sd(y),
    fun.min = function(y) mean(y) - sd(y),
    size = .8
  ) +
  stat_summary(
    geom = "text",
    aes(
      color = day_night,
      label = paste0(sprintf("%2.1f", stat(y) / 1000), "K")
    ),
    position = position_dodge(width = .6),
    hjust = -.2, size = 5.5, family = "Tabular", fontface = "bold"
  ) +
  coord_cartesian(clip = "off") +
  scale_x_discrete(
    labels = str_to_title
  ) +
  scale_y_continuous(
    labels = scales::comma_format(), 
    expand = c(0, 0), 
    limits = c(0, NA)
  ) +
  scale_color_manual(values = colors) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Reported bike shares in London during <span style='color:", colors[1], ";'>day</span> and <span style='color:", colors[2], ";'>night</span> times"),
    subtitle = "TfL bike sharing data from 2015 to 2016 per season and time of day.\nErrorbars show the mean ± standard deviation."
  ) +
  theme_minimal(base_size = 18, base_family = "Cabinet Grotesk") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title.position = "plot",
    plot.title = ggtext::element_markdown(face = "bold", size = 26),
    plot.subtitle = element_text(color = "grey30", margin = margin(t = 6, b = 12)),
    axis.text.x = element_text(size = 17, face = "bold"),
    axis.text.y = element_text(family = "Tabular"),
    axis.line.x = element_line(size = 1.2, color = "grey65"),
    plot.margin = margin(rep(15, 4))
  )
```

## Import the Data Set

```{r}
#| label: import-data-exercise
library(tidyverse)

bikes <- readr::read_csv(
  "https://raw.githubusercontent.com/z3tt/graphic-design-ggplot2/main/data/london-bikes-custom.csv", 
  col_types = "Dcfffilllddddc"
)

bikes$season <- forcats::fct_inorder(bikes$season)
```

## Create Sina Plot

```{r}
#| label: colors-exercise-ex-1-1
#| eval: false
#| code-line-numbers: "5,6,7,8,9|7"
ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  ggforce::geom_sina(
    aes(color = day_night),
    position = position_dodge(width = .6),
    alpha = .5
  ) +
  theme_minimal(
    base_size = 18, 
    base_family = "Cabinet Grotesk"
  )
```

## Create Sina Plot

```{r}
#| label: colors-exercise-ex-1-1
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Add Errorbars

```{r}
#| label: colors-exercise-ex-1-2
#| eval: false
#| code-line-numbers: "10,11,12,13,14|6,11|7,12"
ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  ggforce::geom_sina(
    aes(color = day_night),
    position = position_dodge(width = .6),
    alpha = .5
  ) + 
  stat_summary(
    aes(color = day_night),
    position = position_dodge(width = .6),
    size = .8
  ) +
  theme_minimal(
    base_size = 18, 
    base_family = "Cabinet Grotesk"
  )
```

## Add Errorbars

```{r}
#| label: colors-exercise-ex-1-2
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Use Lighter Point Colors

```{r}
#| label: colors-exercise-ex-1-3
#| eval: false
#| code-line-numbers: "6,7,8,9"
ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  ggforce::geom_sina(
    aes(color = stage(
      day_night, 
      after_scale = lighten(color, .6)
    )),
    position = position_dodge(width = .6),
    alpha = .5
  ) + 
  stat_summary(
    aes(color = day_night),
    position = position_dodge(width = .6),
    size = .8
  ) +
  theme_minimal(
    base_size = 18, 
    base_family = "Cabinet Grotesk"
  )
```

## Use Lighter Point Colors

```{r}
#| label: colors-exercise-ex-1-3
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Use Standard Deviation

```{r}
#| label: colors-exercise-ex-1-4
#| eval: false
#| code-line-numbers: "15,16,17"
p1 <- ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  ggforce::geom_sina(
    aes(color = stage(
      day_night, 
      after_scale = lighten(color, .6)
    )),
    position = position_dodge(width = .6),
    alpha = .5
  ) + 
  stat_summary(
    aes(color = day_night),
    fun = mean,
    fun.max = function(y) mean(y) + sd(y),
    fun.min = function(y) mean(y) - sd(y),
    position = position_dodge(width = .6),
    size = .8
  ) +
  theme_minimal(
    base_size = 18, 
    base_family = "Cabinet Grotesk"
  )

p1
```

## Add Annotations

```{r}
#| label: colors-exercise-ex-1-4
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Add Annotations

```{r}
#| label: colors-exercise-ex-1-5
#| eval: false
#| code-line-numbers: "2,3,4,5,6,7,8,9,10,11,12,13|3|6,7,8|10"
p2 <- p1 +
  stat_summary(
    geom = "text",
    aes(
      color = day_night,
      label = paste0(
        sprintf("%2.1f", stat(y) / 1000), "K"
      )
    ),
    position = position_dodge(width = .6),
    hjust = -.2, family = "Tabular", 
    size = 5.5, fontface = "bold"
  ) 

p2
```

## Add Annotations

```{r}
#| label: colors-exercise-ex-1-5
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Adjust Axes + Clipping

```{r}
#| label: colors-exercise-ex-1-6
#| eval: false
#| code-line-numbers: "2|3,4,5|6,7,8,9,10"
p3 <- p2 +
  coord_cartesian(clip = "off") +
  scale_x_discrete(
    labels = str_to_title
  ) +
  scale_y_continuous(
    labels = scales::comma_format(), 
    expand = c(0, 0), 
    limits = c(0, NA)
  ) 

p3
```

## Adjust Axes + Clipping

```{r}
#| label: colors-exercise-ex-1-6
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Add Colors + Labels

```{r}
#| label: colors-exercise-ex-1-7
#| eval: false
#| code-line-numbers: "1|4,5,6|7,8,9,10,11"
colors <- c("#987708", "#663399")

p4 <- p3  +
  scale_color_manual(
    values = colors
  ) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Reported bike shares in London during <span style='color:", colors[1], ";'>day</span> and <span style='color:", colors[2], ";'>night</span> times"),
    subtitle = "TfL bike sharing data from 2015 to 2016 per season and time of day.\nErrorbars show the mean ± standard deviation."
  )

p4
```

## Add Colors + Labels

```{r}
#| label: colors-exercise-ex-1-7
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Theme Styling

```{r}
#| label: colors-exercise-ex-1-8
#| eval: false
#| code-line-numbers: "2|3,4|5,6,7|8,9|10|11"
p4 +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title.position = "plot",
    plot.title = ggtext::element_markdown(face = "bold", size = 26),
    plot.subtitle = element_text(color = "grey30", margin = margin(t = 6, b = 12)),
    axis.text.x = element_text(size = 17, face = "bold"),
    axis.text.y = element_text(family = "Tabular"),
    axis.line.x = element_line(size = 1.2, color = "grey65"),
    plot.margin = margin(rep(15, 4))
  )
```

## Theme Styling

```{r}
#| label: colors-exercise-ex-1-8
#| echo: false
#| fig-width: 10
#| fig-align: "center"
```

## Full Code

```{r}
#| label: exercise-jittered-errorbars-full-code
#| eval: false
library(tidyverse)
library(colorspace)
library(ggtext)

bikes <- readr::read_csv(
  "https://raw.githubusercontent.com/z3tt/graphic-design-ggplot2/main/data/london-bikes-custom.csv", 
  col_types = "Dcfffilllddddc"
)

bikes$season <- forcats::fct_inorder(bikes$season)

colors <- c("#987708", "#663399")

ggplot(bikes, aes(x = season, y = count)) +
  ggforce::geom_sina(
    aes(
      color = stage(
        day_night, after_scale = lighten(color, .6)
    )),
    position = position_dodge(width = .6),
    alpha = .5
  ) + 
  stat_summary(
    aes(color = day_night),
    position = position_dodge(width = .6),
    fun = mean,
    fun.max = function(y) mean(y) + sd(y),
    fun.min = function(y) mean(y) - sd(y),
    size = .8
  ) +
  stat_summary(
    geom = "text",
    aes(
      color = day_night,
      label = paste0(sprintf("%2.1f", stat(y) / 1000), "K")
    ),
    position = position_dodge(width = .6),
    hjust = -.2, size = 5.5, family = "Tabular", fontface = "bold"
  ) +
  coord_cartesian(clip = "off") +
  scale_x_discrete(
    labels = str_to_title
  ) +
  scale_y_continuous(
    labels = scales::comma_format(), 
    expand = c(0, 0), 
    limits = c(0, NA)
  ) +
  scale_color_manual(values = colors) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Reported bike shares in London during <span style='color:", colors[1], ";'>day</span> and <span style='color:", colors[2], ";'>night</span> times"),
    subtitle = "TfL bike sharing data from 2015 to 2016 per season and time of day.\nErrorbars show the mean ± standard deviation."
  ) +
  theme_minimal(base_size = 18, base_family = "Cabinet Grotesk") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title.position = "plot",
    plot.title = element_markdown(face = "bold", size = 26),
    plot.subtitle = element_text(color = "grey30", margin = margin(t = 6, b = 12)),
    axis.text.x = element_text(size = 17, face = "bold"),
    axis.text.y = element_text(family = "Tabular"),
    axis.line.x = element_line(size = 1.2, color = "grey65"),
    plot.margin = margin(rep(15, 4))
  )
```
