---
title: "Graphic Design with ggplot2"
subtitle: "<span style='font-size:1.75em;line-height:1.1;'>Working with Labels and Annotations</span></span>"
author: "<b style='color:#28a87d;'>Cédric Scherer</b> **//** rstudio::conf **//** July 2022"
format: 
  revealjs: 
    theme: css/style.scss
    footer: '**[Cédric Scherer](https::/cedricscherer.com) //** rstudio::conf **//** July 2022'
    #'rstudio::conf(workshop = "Graphic Design with ggplot2", instructor = "Cédric Scherer")'
    transition: fade
    progress: true
    multiplex: false
    preview-links: true
    scrollable: false
    hide-inactive-cursor: true
    highlight-style: printing
    chalkboard:
      theme: whiteboard
      boardmarker-width: 4
      buttons: true
    pause: true
    mainfont: Cabinet Grotesk
editor: source
knitr:
  opts_chunk: 
    dev: "ragg_png"
    retina: 1
    dpi: 300
execute:
  freeze: auto
  echo: true
  fig-width: 5
  fig-height: 6
---

## Setup

```{r}
#| label: prep
library(tidyverse)

bikes <- readr::read_csv(
  here::here("data", "london-bikes-custom.csv"), 
  col_types = "Dcfffilllddddc"
)

#bikes$season <- factor(bikes$season, levels = c("spring", "summer", "autumn", "winter"))
bikes$season <- forcats::fct_inorder(bikes$season)

theme_set(theme_light(base_size = 14, base_family = "Roboto Condensed"))

theme_update(
  panel.grid.minor = element_blank(),
  legend.position = "top"
)
```

# Labels + `theme()`

## Working with Labels

```{r}
#| label: ggplot-labs-bundled-extended
#| output-location: column
#| code-line-numbers: "9,10,11,12,13,14,15,16,17"
g <- ggplot(
    bikes, 
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  ) +
  labs(
    x = "Feels-like temperature (°F)",
    y = "Reported bike shares",
    title = "TfL bike sharing trends",
    subtitle = "Reported bike rents versus feels-like temperature in London",
    caption = "Data: TfL",
    color = "Season:",
    tag = "1."
  )

g
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-theme-title
#| output-location: column
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot"
)
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-labs-theme-element-text
#| output-location: column
#| code-line-numbers: "4,5,6"
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot",
  axis.text = element_text(
    color = "grey40"
  )
)
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-labs-theme-element-text-more
#| output-location: column
#| code-line-numbers: "6|7|8|9|10|11|12"
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot",
  axis.text = element_text(
    color = "grey40", 
    family = "Tabular",
    face = "italic",
    hjust = 1,
    vjust = 0,
    angle = 45,
    lineheight = 1.3, ## no effect here
    margin = margin(10, 0, 20, 0)
  )
)
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-labs-theme-element-text-x
#| output-location: column
#| code-line-numbers: "16,17,18"
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot",
  axis.text = element_text(
    color = "grey40", 
    family = "Tabular",
    face = "italic",
    colour = NULL,
    size = NULL,
    hjust = 1,
    vjust = 0,
    angle = 45,
    lineheight = 1.3, ## no effect here
    margin = margin(10, 0, 20, 0) ## no effect here
  ),
  axis.text.x = element_text(
    margin = margin(10, 0, 20, 0) ## trbl
  )
)
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-labs-theme-element-text-tag
#| output-location: column
#| code-line-numbers: "16,17,18"
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot",
  axis.text = element_text(
    color = "grey40", 
    family = "Tabular",
    face = "italic",
    colour = NULL,
    size = NULL,
    hjust = 1,
    vjust = 0,
    angle = 45,
    lineheight = 1.3, ## no effect here
    margin = margin(10, 0, 20, 0) ## no effect here
  ),
  plot.tag = element_text(
    margin = margin(0, 12, -8, 0) ## trbl
  )
)
```

## Customize Labels via \`theme()\`

```{r}
#| label: ggplot-labs-theme-element-text-debug
#| output-location: column
#| code-line-numbers: "15,19"
g + theme(
  plot.title = element_text(face = "bold"),
  plot.title.position = "plot",
  axis.text = element_text(
    color = "grey40", 
    family = "Tabular",
    face = "italic",
    colour = NULL,
    size = NULL,
    hjust = 1,
    vjust = 0,
    angle = 45,
    lineheight = 1.3, ## no effect here
    margin = margin(10, 0, 20, 0), ## no effect here
    debug = TRUE
  ),
  plot.tag = element_text(
    margin = margin(0, 12, -8, 0), ## trbl
    debug = TRUE
  )
)
```

# Labels + `scale_*()`

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-scales-prep
#| output-location: column
#| code-line-numbers: "1|3,4,5,6"
g <- g + labs(title = NULL, subtitle = NULL, tag = NULL) 

g +
  scale_y_continuous(
    breaks = 0:4*15000
  ) 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-scales-1
#| output-location: column
#| code-line-numbers: "4"
g +
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = scales::comma_format()
  ) 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-scales-2
#| output-location: column
#| code-line-numbers: "5|7"
g + 
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = scales::comma_format(
      suffix = " bikes"
    ),
    name = NULL
  ) 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-scales-3
#| output-location: column
#| code-line-numbers: "5|9,10,11,12,13"
g + 
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = scales::comma_format(
      suffix = "\nbikes shared"
    ),
    name = NULL
  ) +
  theme(
    axis.text.y = element_text(
      hjust = .5
    )
  )
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-scales-4
#| output-location: column
#| code-line-numbers: "5|5,7"
g +
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = scales::comma_format(
      scale = .001
    ),
    name = "Reported bike shares in thousands"
  )
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-function
#| output-location: column
#| code-line-numbers: "4"
g +
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = function(y) y / 1000,
    name = "Reported bike shares in thousands",
  )
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-continuous-labels-function-paste
#| output-location: column
#| code-line-numbers: "2|3|4"
g +
  scale_x_continuous(
    labels = function(y) paste0(y, "°F"),
    name = "Feels-Like Temperature"
  ) 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-discrete-labels
#| output-location: column
ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  geom_boxplot() +
  scale_x_discrete() 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-xy-discrete-labels-function
#| output-location: column
#| code-line-numbers: "8|7,8"
ggplot(
    bikes, 
    aes(x = season, y = count)
  ) +
  geom_boxplot() +
  scale_x_discrete(
    name = NULL,
    labels = stringr::str_to_title
  ) 
```

## Format Labels via \`scale\_\*\`

```{r}
#| label: scales-color-discrete-name-labels
#| output-location: column
#| code-line-numbers: "2,4"
g +
  scale_color_discrete(
    name = NULL,
    labels = stringr::str_to_title
  )
```

# Labels +<br>`ggtext::element_markdown`

## Styling Labels with {ggtext}

```{r}
#| label: ggtext-prep-markdown
#| output-location: column
#| code-line-numbers: "2"

g + 
  ggtitle("**TfL bike sharing trends by _season_**")
```

## Styling Labels with {ggtext}

```{r}
#| label: ggtext-element-markdown
#| output-location: column
#| code-line-numbers: "4,5,6,7|4|6"
# install.packages("ggtext")

g + 
  ggtitle("**TfL bike sharing trends by _season_**") +
  theme(
    plot.title = ggtext::element_markdown()
  )
```

## Styling Labels with {ggtext}

```{r}
#| label: ggtext-element-markdown-html
#| output-location: column
#| code-line-numbers: "4|4,6"
#| fig-height: 5.5
# install.packages("ggtext")

g + 
  ggtitle("<b style='font-family:tabular;font-size:25pt'>TfL</b> bike sharing trends by <i style='color:#28a87d;'>season</i>") +
  theme(
    plot.title = ggtext::element_markdown()
  )
```

<p style="font-size:12pt;">

`<b style='font-family:tabular;font-size:25pt'>TfL</b> bike sharing trends by <i style='color:#28a87d;'>season</i>`

</p>

# Labels + `facet_*()`

## Facet Labellers

```{r}
#| label: facet-wrap-labels-1
#| output-location: column
#| code-line-numbers: "4"
g +
  facet_wrap(
    ~ day_night,
    labeller = label_both
  )
```

## Facet Labellers

```{r}
#| label: facet-wrap-labels-2
#| output-location: column
#| code-line-numbers: "3,4"
g +
  facet_wrap(
    ~ is_workday + day_night,
    labeller = label_both
  )
```

## Facet Labellers

```{r}
#| label: facet-wrap-labels-3
#| output-location: column
#| code-line-numbers: "4,6|4,5,6"
g +
  facet_wrap(
    ~ is_workday + day_night,
    labeller = labeller(
      day_night = stringr::str_to_title
    )
  )
```

## Facet Labellers

```{r}
#| label: facet-wrap-labels-4
#| output-location: column
#| code-line-numbers: "1,2,3,4|10,11"
codes <- c(
  `TRUE` = "Workday", 
  `FALSE` = "Weekend or Holiday"
)

g +
  facet_wrap(
    ~ is_workday + day_night,
    labeller = labeller(
      day_night = stringr::str_to_title,
      is_workday = codes
    )
  )
```

## Facet Labellers

```{r}
#| label: facet-wrap-labels-5
#| output-location: column
#| code-line-numbers: "10"
codes <- c(
  `TRUE` = "Workday", 
  `FALSE` = "Weekend or Holiday"
)

g +
  facet_wrap(
    ~ is_workday + day_night,
    labeller = labeller(
      .default = stringr::str_to_title,
      is_workday = codes
    )
  )
```

## Facet Labeller

```{r}
#| label: facet-grid-labeller
#| output-location: column
#| code-line-numbers: "6,7,8,9"
g +
  facet_grid(
    day_night ~ is_workday,
    scales = "free",
    space = "free",
    labeller = labeller(
      day_night = stringr::str_to_title,
      is_workday = codes
    )
  )
```

# Handling Long Labels

## Handling Long Labels with {stringr}

```{r}
#| label: stringr-default-labs
#| output-location: column
ggplot(
    bikes, 
    aes(x = weather_type, 
        y = count)
  ) +
  geom_boxplot()
```

```{r}
#| label: stringr-wrap-labs
#| output-location: column
#| code-line-numbers: "3"
ggplot(
    bikes, 
    aes(x = stringr::str_wrap(weather_type, 6), 
        y = count)
  ) +
  geom_boxplot()
```

## Handling Long Labels with {ggtext}

```{r}
#| label: ggtext-text-default
#| output-location: column
#| code-line-numbers: "2|3,4,5,6|4"
g + 
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = element_text(size = 20),
    plot.title.position = "plot"
  )
```

## Handling Long Labels

```{r}
#| label: ggtext-textbox-default
#| output-location: column
#| code-line-numbers: "2|3,4,5,6|4"
g + 
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = 
      ggtext::element_textbox_simple(size = 20),
    plot.title.position = "plot"
  )
```

## Handling Long Labels

```{r}
#| label: ggtext-textbox-margin
#| output-location: column
#| code-line-numbers: "5,6"
g + 
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = ggtext::element_textbox_simple(
      margin = margin(t = 12, b = 12),
      lineheight = .9
    ),
    plot.title.position = "plot"
  )
```

## Handling Long Labels

```{r}
#| label: ggtext-textbox-box
#| output-location: column
#| code-line-numbers: "6"
g + 
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = ggtext::element_textbox_simple(
      margin = margin(t = 12, b = 12),
      fill = "grey90",
      lineheight = .9
    ),
    plot.title.position = "plot"
  )
```

## Handling Long Labels

```{r}
#| label: ggtext-textbox-box-styled
#| output-location: column
#| code-line-numbers: "6|8|9|10|11"
g + 
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = ggtext::element_textbox_simple(
      margin = margin(t = 12, b = 12),
      padding = margin(rep(12, 4)),
      fill = "grey90",
      box.color = "grey40",
      r = unit(9, "pt"),
      halign = .5,
      face = "bold",
      lineheight = .9
    ),
    plot.title.position = "plot"
  )
```

# Annotations with `annotate()`

## Add Single Text Annotations

```{r}
#| label: annotate-text
#| output-location: column
#| code-line-numbers: "3,4,5,6,7,8|4|5,6|7"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext"
  )
```

## Style Text Annotations

```{r}
#| label: annotate-text-adj
#| output-location: column
#| code-line-numbers: "8,9,10,11"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    color = "firebrick",
    fontface = "bold",
    lineheight = .9
  )
```

## Add Multiple Text Annotations

```{r}
#| label: annotate-text-multiple
#| output-location: column
#| code-line-numbers: "8,9,10,11"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = c(90, 50),
    y = c(27.5, 3.5),
    label = c("Text A", "Text B"),
    color = c("black", "firebrick"),
    size = c(5, 10),
    fontface = c("plain", "bold")
  )
```

## Add Boxes (Rectangles)

```{r}
#| label: annotate-box
#| output-location: column
#| code-line-numbers: "4,5,6,7,8,9|4|5,6,7,8"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "rect",
    xmin = -Inf, 
    xmax = 60, 
    ymin = 20, 
    ymax = Inf,
    fill = "#663399"  
  )
```

## Add Boxes (Rectangles)

```{r}
#| label: annotate-box-hierarchy
#| output-location: column
#| code-line-numbers: "1,2,9,10"
ggplot(bikes, aes(humidity, temp)) +
  annotate(
    geom = "rect",
    xmin = -Inf, 
    xmax = 60, 
    ymin = 20, 
    ymax = Inf,
    fill = "#663399"
  ) +
  geom_point(size = 2, color = "grey")
```

## Add Lines (Segments)

```{r}
#| label: annotate-text-line
#| output-location: column
#| code-line-numbers: "11,12,13,14|12|13|14"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    lineheight = .9 
  ) +
  annotate(
    geom = "segment",
    x = 90, xend = 82,
    y = 25, yend = 18.5
  )
```

## Add Curves

```{r}
#| label: annotate-text-curve
#| output-location: column
#| code-line-numbers: "12"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    lineheight = .9 
  ) +
  annotate(
    geom = "curve",
    x = 90, xend = 82,
    y = 25, yend = 18.5
  ) 
```

## Add Arrows

```{r}
#| label: annotate-text-arrow
#| output-location: column
#| code-line-numbers: "13"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate( 
    geom = "text", 
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    lineheight = .9 
  ) +
  annotate(
    geom = "curve",
    x = 90, xend = 82,
    y = 25, yend = 18.5,
    curvature = -.3,
    arrow = arrow()
  )
```

## Add Arrows

```{r}
#| label: annotate-text-arrow-type2
#| output-location: column
#| code-line-numbers: "12,14,15,16,17|14,15,16"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate( 
    geom = "text", 
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    lineheight = .9 
  ) +
  annotate(
    geom = "curve",
    x = 90, xend = 82,
    y = 25, yend = 18.5,
    curvature = -.3,
    arrow = arrow(
      length = unit(10, "pt"),
      type = "closed",
      ends = "both"
    )
  )
```

## Add Arrows

```{r}
#| label: annotate-text-arrow-skewed
#| output-location: column
#| code-line-numbers: "15,16"
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate( 
    geom = "text", 
    x = 90,
    y = 27.5,
    label = "Some\nadditional\ntext",
    size = 6,
    lineheight = .9 
  ) +
  annotate(
    geom = "curve",
    x = 94, xend = 82,
    y = 26, yend = 18.5,
    curvature = -.8,
    angle = 140,
    arrow = arrow(
      length = unit(10, "pt"),
      type = "closed"
    )
  )
```

# Annotations with \`geom\_\*()\`'s'

## Highlight Hot Periods

```{r}
#| label: highlight-hot-periods
#| output-location: column
#| code-line-numbers: "1,2,3,4,5,6,7,8,9|2|2,9|5,6,7,8|6"
g <- ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5)
```

## Annotations with \`geom_text()\`

```{r}
#| label: highlight-geom-text
#| output-location: column
#| code-line-numbers: "10,11,12,13,14"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  geom_text(
    aes(label = season),
    nudge_x = .3,
    hjust = 0
  ) 
```

## Annotations with \`geom_label()\`

```{r}
#| label: highlight-geom-label
#| output-location: column
#| code-line-numbers: "10"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  geom_label(
    aes(label = season),
    nudge_x = .3,
    hjust = 0
  ) 
```

## Annotations with {ggrepel}

```{r}
#| label: seeding
#| include: false
set.seed(20220726)
```

```{r}
#| label: highlight-geom-text-repel
#| output-location: column
#| code-line-numbers: "10,11,12"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = season)
  )
```

## Annotations with {ggrepel}

```{r}
#| label: highlight-geom-text-repel-color
#| output-location: column
#| code-line-numbers: "4,14,15,16,17"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = str_to_title(season))
  ) +
  scale_color_manual(
    values = c("firebrick", "black"), 
    guide = "none"
  )
```

## Annotations with {ggrepel}

```{r}
#| label: highlight-geom-text-repel-dist
#| output-location: column
#| code-line-numbers: "13,14,15,16"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = str_to_title(season)),
    ## space between points + labels
    box.padding = .4,
    ## always draw segments
    min.segment.length = 0
  ) +
  scale_color_manual(
    values = c("firebrick", "black"), 
    guide = "none"
  )
```

## Annotations with {ggrepel}

```{r}
#| label: highlight-geom-text-repel-lim
#| output-location: column
#| code-line-numbers: "13,14,20"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = str_to_title(season)),
    ## force to the right
    xlim = c(NA, 35), hjust = 1
  ) +
  scale_color_manual(
    values = c("firebrick", "black"), 
    guide = "none"
  ) + 
  xlim(25, NA)
```

## Annotations with {ggrepel}

```{r}
#| label: highlight-geom-text-repel-arrow
#| output-location: column
#| code-line-numbers: "15,16,17"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = str_to_title(season)),
    ## force to the right
    xlim = c(NA, 35), 
    ## style segment
    segment.curvature = .01,
    arrow = arrow(length = unit(.02, "npc"), type = "closed")
  ) +
  scale_color_manual(
    values = c("firebrick", "black"), 
    guide = "none"
  ) + 
  xlim(25, NA)
```

## Annotations with {ggrepel}

```{r}
#| label: highlight-geom-text-repel-inflect
#| output-location: column
#| code-line-numbers: "15,16,17"
ggplot(
    filter(bikes, temp >= 27),
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +
  ggrepel::geom_text_repel(
    aes(label = str_to_title(season)),
    ## force to the right
    xlim = c(NA, 35), 
    ## style segment
    segment.curvature = .001,
    segment.inflect = TRUE
  ) +
  scale_color_manual(
    values = c("firebrick", "black"), 
    guide = "none"
  ) + 
  xlim(25, NA)
```

## Annotations with {ggforce}

```{r}
#| label: highlight-geom-mark-prep
#| output-location: column
#| code-line-numbers: "1,2,3,4,5,6,7,8,9,10,11,12,13,14|2|4,11,12,13,14"
ggplot(
    filter(bikes, temp > 20 & season != "summer"),
    aes(x = humidity, y = temp, 
        color = season)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point() +
  scale_color_brewer(
    palette = "Dark2",
    guide = "none"
  ) 
```

## Annotations with {ggforce}

```{r}
#| label: highlight-geom-mark-rect
#| output-location: column
#| code-line-numbers: "11,12,13"
ggplot(
    filter(bikes, temp > 20 & season != "summer"),
    aes(x = humidity, y = temp, 
        color = season)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point() +
  ggforce::geom_mark_rect(
    aes(label = str_to_title(season))
  ) + 
  scale_color_brewer(
    palette = "Dark2",
    guide = "none"
  ) 
```

## Annotations with {ggforce}

```{r}
#| label: highlight-geom-mark-limit
#| output-location: column
#| code-line-numbers: "18"
ggplot(
    filter(bikes, temp > 20 & season != "summer"),
    aes(x = humidity, y = temp,
        color = season)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point() +
  ggforce::geom_mark_rect(
    aes(label = str_to_title(season))
  ) + 
  scale_color_brewer(
    palette = "Dark2",
    guide = "none"
  ) +
  ylim(NA, 35)
```

## Annotations with {ggforce}

```{r}
#| label: highlight-geom-mark-customize
#| output-location: column
#| code-line-numbers: "13,14,15,16,17,18|13,14|15,16,17|18"
ggplot(
    filter(bikes, temp > 20 & season != "summer"),
    aes(x = humidity, y = temp, 
        color = season)
  ) +
  geom_point(
    data = bikes, 
    color = "grey65", alpha = .3
  ) +
  geom_point() +
  ggforce::geom_mark_rect(
    aes(label = str_to_title(season)),
    expand = unit(5, "pt"),
    radius = unit(0, "pt"),
    con.cap = unit(0, "pt"),
    label.buffer = unit(15, "pt"),
    con.type = "straight",
    label.fill = "transparent"
  ) + 
  scale_color_brewer(
    palette = "Dark2",
    guide = "none"
  ) +
  ylim(NA, 35)
```

## Annotations with {ggforce}

```{r}
#| label: highlight-geom-mark-filter
#| output-location: column
#| code-line-numbers: "1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16|1,2,3,4,5|4,13,14,15,16|7|8,9|10"
ggplot(
    bikes,
    aes(x = humidity, y = temp, 
        color = season == "summer")
  ) +
  geom_point(alpha = .4) +
  ggforce::geom_mark_hull(
    aes(label = str_to_title(season),
        filter = season == "summer",
        description = "June to August"),
    expand = unit(10, "pt")    
  ) + 
  scale_color_manual(
    values = c("grey65", "firebrick"),
    guide = "none"
  )
```

# Adding Images

## Load and Modify Image

```{r}
#| label: read-img
#| code-line-numbers: "1,2,3,4,5|1|2|3"
url <- "https://d33wubrfki0l68.cloudfront.net/dbb07b06a7b3fe056db386fef0b158cc2fd33cb9/8b491/assets/img/2022conf/logo-rstudio-conf.png"
img <- magick::image_read(url)
img <- magick::image_negate(img)

img
```

## Add Background Image

```{r}
#| label: annotate-custom
#| output-location: column
#| code-line-numbers: "3,4,5,6|3|4"
ggplot(bikes, aes(date, temp_feel)) +
  annotation_custom(
    grid::rasterGrob(
      image = img
    ) 
  ) +
  geom_point(color = "#71a5d4")
```

## Adjust Position

```{r}
#| label: annotate-custom-position
#| output-location: column
#| code-line-numbers: "5,6,7,11"
ggplot(bikes, aes(date, temp_feel)) +
  annotation_custom(
    grid::rasterGrob(
      image = img,
      x = .5,
      y = .9,
      width = .9 
    ) 
  ) +
  geom_point(color = "#71a5d4") +
  ylim(NA, 37)
```

## Place Image Outside

```{r}
#| label: annotate-custom-position-outside
#| output-location: column
#| code-line-numbers: "5,6,11,12,13,14|6,11,13"
ggplot(bikes, aes(date, temp_feel)) +
  annotation_custom(
    grid::rasterGrob(
      image = img,
      x = .47,
      y = 1.15,
      width = .9 
    ) 
  ) +
  geom_point(color = "#71a5d4") +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(90, 10, 10, 10)
  )
```

## <b style='color:#663399;'>Recap</b>

::: recap
::: incremental
-   style labels such as title, axis and legend texts with **`theme()`**
-   format data-related labels with the **`labels`** argument of `scale_*()`
-   adjust strip text with the **`facet_*(labeller)`** functionality
-   add data-related annotations with **`geom_text|label()`**
-   ... and data-unrelated annotations with **`annotate()`**
-   **`annotation_custom(rasterGrob())`** is a basic way to add images
-   **`{ggtext}`** allows to render labels with markdown and basic html
-   `{ggtext}` also allows to add **dynamic linebreaks** and images
-   **`{ggrepel}`** ensures clever placement of annotations
-   **`ggforce::geom_mark_*()`** provide a set of advanced annotations
:::
:::

# Exercises

## Exercise 1

::: question
-   {ggtext} also comes with some new geom's. Explore those and other options on the package webpage: [wilkelab.rg/ggtext](https://wilkelab.org/ggtext/).
-   **Create the following visualization, as close as possible**, with the `penguins` dataset which is provided by the [{palmerpenguins} package](https://allisonhorst.github.io/palmerpenguins/).
    -   For the species labels, you likely have to create a summary data set.
    -   Use the {ggtext} geometries and theme elements to format the labels.
    -   Also, make use of the other components such as scales, original theme, and theme customization.
:::

------------------------------------------------------------------------

```{r}
#| label: exercise-text-styling
#| fig-width: 6
#| fig-height: 3.7
#| fig-pos: "center"
#| echo: false
#| purl: false
library(palmerpenguins)

penguins_labs <-
  penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(
    species_lab = case_when(
      species == "Adelie" ~ "<b style='font-size:15pt;'>*P. adéliae*</b><br>(Adélie penguin)",
      species == "Chinstrap" ~ "<b style='font-size:15pt;'>*P. antarctica*</b><br>(Chinstrap penguin)",
      species == "Gentoo" ~ "<b style='font-size:15pt;'>*P. papua*</b><br>(Gentoo penguin)"
    )
  )

ggplot(
    penguins, 
    aes(x = bill_length_mm, y = bill_depth_mm, 
        color = species, size = body_mass_g)
  ) +
  geom_point(alpha = .2, stroke = .3) +
  geom_point(shape = 1, stroke = .3) +
  ggtext::geom_richtext(
    data = penguins_labs,
    aes(label = species_lab, color = species, 
        color = after_scale(colorspace::darken(color, .4))),
    family = "Roboto Condensed",
    size = 3, lineheight = .8,
    fill = "#ffffffab", ## hex-alpha code
    show.legend = FALSE
  ) +
  coord_cartesian(
    expand = FALSE,
    clip = "off"
  )  +
  scale_x_continuous(
    limits = c(30, 60),
    breaks = 6:12*5
  ) +
  scale_y_continuous(
    limits = c(12.5, 22.5),
    breaks = seq(12.5, 22.5, by = 2.5)
  ) +
  scale_color_manual(
    guide = "none",
    values = c("#FF8C00", "#A034F0", "#159090")
  ) +
  scale_size(
    name = "Body mass:",
    breaks = 3:6 * 1000,
    labels = function(x) paste(x / 1000, "kg"),
    range = c(.25, 4.5)
  ) +
  labs(
    x = "Bill length *(mm)*",
    y = "Bill depth *(mm)*",
    title = "Bill dimensions of brush-tailed penguins *Pygoscelis spec.*",
    caption = "Horst AM, Hill AP, Gorman KB (2020). <span style='font-family:tabular;'>palmerpenguins</span> R package version 0.1.0"
  ) +
  theme_minimal(
    base_size = 10, base_family = "Roboto Condensed"
  ) +
  theme(
    plot.title = ggtext::element_markdown(
      face = "bold", size = 16, margin = margin(12, 0, 12, 0)
    ),
    plot.title.position = "plot",
    plot.caption = ggtext::element_markdown(
      size = 7, color = "grey50",
      margin = margin(12, 0, 6, 0)
    ),
    plot.caption.position = "plot",
    axis.text = element_text(family = "Tabular"),
    axis.title.x = ggtext::element_markdown(margin = margin(t = 8)),
    axis.title.y = ggtext::element_markdown(margin = margin(r = 8)),
    panel.grid.minor = element_blank(),
    legend.text = element_text(color = "grey50"),
    plot.margin = margin(0, 14, 0, 12),
    plot.background = element_rect(fill = NA, color = "grey50", size = 1)
  )
```

```{r}
#| label: countdown-ex-1
#| echo: false
countdown::countdown(minutes = 20, font_size = ".5em")
```

## Exercise 2

::: question
-   **Create this logo with the image file `exercises/`:**
:::

<br>

```{r}
#| label: annotate-custom-rstudioconf
#| echo: false
#| fig-align: "center"
ggplot(mapping = aes(x = 0, y = 0)) +
  annotation_custom(
    grid::rasterGrob(
      image = img,
      x = .5,
      y = .1
    ) 
  ) +
  geom_point(
    color = "#71a5d4", size = 150
  ) +
  geom_text(
    label = "R", size = 80,
    family = "Tabular", 
    color = "white"
  ) +
  xlim(-5, 5) +
  ylim(-7.5, 5) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = "#fffdf9", color = "#fffdf9"
    )
  )
```

```{r}
#| label: countdown-ex-2
#| echo: false
countdown::countdown(minutes = 10, font_size = ".5em")
```
